Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-30 ba_resubmit_template_files/main.tex
fix typos using antidote 11.

In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and describes the loss function used to calculate the Bayes estimate for the bipartite matching. Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}. Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study. Finally, Section \ref{discussion} summarizes our contributions and highlights areas for further research.

@@

\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.  He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$, and replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\bm{Z}$ that he calls the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.  Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, where $\alpha_{\pi}$ and $\beta_{\pi}$ are known hyperparameters. It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A ! / (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely. Thus, the prior distribution used by \cite{sadinle_bayesian_2017} is given by
\begin{align}
\label{eqn:sadinle_prior}
p(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$. This makes the sampler inherently sequential, which can be slow when working on linkage tasks with more than a few thousand records.

@@

After computing all comparison vectors within a block, the modeler can further reduce the number of comparison vectors used in the linkage algorithm through indexing. For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields. However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the classical \cite{fellegi_theory_1969} model, leaving the effect of indexing on Bayesian record linkage models to future work.

@@

All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$. These records have the same probability to be identified as the link for record $B_j$. Thus, records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate. Rather than store all of these record labels, we store only a small number $S$. For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form SEI$(r_{p_j}^{ab})$. We collect these memory reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$.

As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$. We provide guidance on choice of $S$ through a simulation in Section \ref{SEI-sensitivity}.

@@

	In Figure \ref{fig:speed2}, where we fix $n_B = 500$, we see near linear growth for the run-time under \texttt{BRL} as $n_A$ increases, and much more static run-time under \texttt{fabl}. The slight increases in the run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data. To illustrate that these trends are generalizable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Supplement \ref{app:appendix-speed}.

@@

	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds. Traceplots do not suggest convergence issues, and are similar to those seen in Supplement \ref{app:appendix-sim} and \ref{app:appendix-es}. As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure has of 9634 matches, with a 95\% credible interval of (9581, 9740). Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.

@@

	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing the comparisons for all $n_A  n_B$ record pairs still can be prohibitive for larger linkage tasks. Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes. Due to the quadratic nature of the comparison space, this computation time would grow quickly with the size of the linkage task, and would be infeasibly slow when dealing with millions of records. Although it is common to use deterministic blocking to reduce the comparison space and then apply probabilistic record linkage within each block, issues arise when sizes of blocks vary across the linkage task. In future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to even larger linkage tasks.

>>>>

In what follows, Section~\ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and describes the loss function used to calculate the Bayes estimate for the bipartite matching. Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}. Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study. Finally, Section~\ref{discussion} summarizes our contributions and highlights areas for further research.

@@

\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.  He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$, and replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\bm{Z}$ that he calls the ``beta distribution for bipartite matching.''  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.  Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, where $\alpha_{\pi}$ and $\beta_{\pi}$ are known hyperparameters. It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A ! / (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely. Thus, the prior distribution used by \cite{sadinle_bayesian_2017} is given by
\begin{align}
\label{eqn:sadinle_prior}
p(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$. This makes the sampler inherently sequential, which can be slow when working on linkage tasks with more than a few thousand records.

@@

After computing all comparison vectors within a block, the modeler can further reduce the number of comparison vectors used in the linkage algorithm through indexing. 
%% antidote 11 flagged as awkward
For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or 
%only 
pairs that exactly match on a specified number of fields. However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the classical \cite{fellegi_theory_1969} model, leaving the effect of indexing on Bayesian record linkage models to future work.

@@

All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$. These records have the same probability to be identified as the link for record $B_j$. Thus, records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate. Rather than store all of these record labels, we store only a small number $S$. For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form SEI$(r_{p_j}^{ab})$. We collect these memory-reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$.

As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$. We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.

@@

	In Figure \ref{fig:speed2}, where we fix $n_B = 500$, we see near linear growth for the run-time under \texttt{BRL} as $n_A$ increases, and much more static run-time under \texttt{fabl}. The slight increases in the run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data. To illustrate that these trends are generalizable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Supplement~\ref{app:appendix-speed}.

@@

	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds. Traceplots do not suggest convergence issues, and are similar to those seen in Supplement \ref{app:appendix-sim} and \ref{app:appendix-es}. As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure has
	%has of 
	9634 matches, with a 95\% credible interval of (9581, 9740). Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.

@@

	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing the comparisons for all $n_A  n_B$ record pairs still can be prohibitive for larger linkage tasks. Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes. Due to the quadratic nature of the comparison space, this computation time would grow quickly with the size of the linkage task, and would be extremely
	%infeasibly 
	slow when dealing with millions of records. Although it is common to use deterministic blocking to reduce the comparison space and then apply probabilistic record linkage within each block, issues arise when sizes of blocks vary across the linkage task. In future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to even larger linkage tasks.
